\section{Conclusions}\label{conclusions}
%Timeline
\paragraph{}
In this project I developed a dynamic monitor for a data flow testing tool, and performed an evaluation to assess its efficiency and its effectiveness. Given the good performances of the tool, I will publish it online to help researchers and practitioners experimenting with data flow criteria --- it could be used for collecting new data on the effectiveness of data flow criteria, or it could be extended with a component that automatically generates test cases for the missing pairs for defining new test case generation approaches. 

\paragraph{}
I was impressed by the low coverage observed on the test suites that I used in the evaluation phase. This suggests that data flow testing is still an open problem that needs new data and techniques, and that provide a complementary view with respect to functional testing and simpler structural criteria like branch and statement testing. 

\paragraph{}
Apart from the results obtained, this project was an invaluable opportunity for me, because it allowed me to apply some concepts I learned during my Bachelor and I had never truly applied before, especially regarding programming principles and design patterns. I learned to write code that can be reused and modularized, and the importance of careful design and planning. Moreover, since I was left a lot of freedom during the development, for the first time I had to learn how to organize the time I could dedicate to the project, and how to use it effectively. The experience I collected from this Bachelor project will be extremely useful in future occasions.

\section{Acknowledgements}
First of all, I would like to thank my advisor Prof. Mauro Pezz\'e for providing me this opportunity. Exploring and working on an unknown topic was a precious experience, and will be very helpful during my future studies. I would also like to thank Mattia Vivanti, which supervised me during the whole project and has always been very helpful and has given me a lot of precious advices. Last but not least, I would like to thank my parents, my girlfriend and my friends for being supportive throughout these three years.


\newpage
\bibliographystyle{abbrv}
\bibliography{references}